---
title: "FinalProject"
author: "Bernardo Santos"
date: "11/3/2021"
output: word_document
---

# Load data set
```{r}
library(dplyr)
library(magrittr)
library(mosaic)
library(rpart)
library(partykit)
library(randomForest)
library(neuralnet)
library(class)
library(caret)
library(mdsr)
library(e1071)

options(scipen = 999)

Funds <- read.csv(file = 'Funds.csv', header = TRUE)
head(Funds, 10)
```

# Check the missing values in the data set
```{r}
na_count <- sapply(Funds, function(y) round(sum(length(which(is.na(y))))/nrow(Funds), digits = 2))
na_count <- data.frame(na_count)
na_count <- cbind(colName = rownames(na_count), na_count)
rownames(na_count) <- 1:nrow(na_count)
colnames(na_count) <- c("colName", "rep_count")
na_count[order(-na_count[,2]),]
```
# Get column names where missing values percentage is greater than 40%
```{r}
na_count_40 <- filter(na_count, rep_count >= 0.4)
na_count_40
```

# Remove the clearly unneeded columns, such as free text and names that do not influence the dataset overall
# Remove quarterly-based returns variables, since we cannot use all of them
```{r}
cols_remove <- c("fund_symbol", "fund_extended_name", "fund_family", "inception_date", "category", "investment_strategy", "currency", "top10_holdings")
cols_remove_join <- c(cols_remove, na_count_40$colName)
q_returns <- c()
for(x in 2012:2020){
  for(y in 1:4){
    if(x == 2020 && y == 4){
      
    } else{
      name <- paste("fund_return_",as.character(x), "_q",as.character(y), sep="")
      if(name %in% cols_remove_join == FALSE){
        q_returns[length(q_returns) + 1] <- name
      }
      
    }
    
  }
}
cols_remove_join <- c(cols_remove_join, q_returns)
new_funds <- Funds %>% 
  dplyr::select(-cols_remove_join)
head(new_funds, 10)
```

# Define target variables
```{r}
# Main targets as numerical variables: fund_return_3months, fund_return_1year, fund_return_5years, fund_return_10years.
# Main target as categorical variables: return_rating.
```

# Perform vectorization
```{r}
encode_ordinal <- function(x, order = unique(x)) {
  x <- as.numeric(factor(x, levels = order, exclude = NULL))
  x
}

new_funds[["investment_type_encoded"]] <- encode_ordinal(new_funds[["investment_type"]])
new_funds[["size_type_encoded"]] <- encode_ordinal(new_funds[["size_type"]])

new_funds <- new_funds %>% 
  dplyr::select(-c("investment_type", "size_type"))
head(new_funds, 10)
```

# Performs standardization
```{r}
# Remove rows that contain missing data.
new_funds <- new_funds[complete.cases(new_funds),]
new_funds

# Convert columns to numeric
cols <- colnames(new_funds)
new_funds[cols] <- sapply(new_funds[cols],as.numeric)
sapply(new_funds, class)

# Max and mins for standardization
maxs <- apply(new_funds[1:102], 2, max) 
mins <- apply(new_funds[1:102], 2, min)

# Perform scaling
funds_scaled <- as.data.frame(scale(new_funds[1:102], center = mins, scale = maxs - mins))
funds_scaled$investment_type_encoded <- new_funds[["investment_type_encoded"]]
funds_scaled$size_type_encoded <- new_funds[["size_type_encoded"]]
head(funds_scaled, 10) # Print scaled data set
```

```{r}
res <- as.data.frame(cor(funds_scaled))
res <- as.data.frame(res["fund_return_10years"])
res <- res %>% 
  filter(abs(res$fund_return_10years) > .3) %>% 
  arrange(desc(fund_return_10years))
res
```

```{r}
# Given the results from the correlation table from the last chunk, 
```


```{r}
funds_scaled <- funds_scaled %>% 
  dplyr::select(rating, return_rating, investment_type_encoded, size_type_encoded, price_earnings_ratio, price_book_ratio, price_sales_ratio, asset_cash, asset_stocks, asset_bonds, sector_basic_materials, sector_utilities, sector_communication_services, sector_technology, fund_return_ytd, fund_return_1year, category_return_1year, fund_return_3years, category_return_3years, fund_return_5years, fund_return_10years, category_return_5years, fund_beta_3years, category_beta_3years, fund_beta_5years, category_beta_5years, fund_mean_annual_return_3years, fund_mean_annual_return_3years, net_asset_value)

head(funds_scaled, 10)
```

# Divide the data into training and testing
```{r}
set.seed(200)

n <- nrow(funds_scaled)
test_idx <- sample.int(n, size = round(0.2 * n))
train <- funds_scaled[-test_idx,]
test <- funds_scaled[test_idx,]
nrow(funds_scaled)
```

# Numerical response variable -> fund_return_10years

```{r}
# The following chunks of code are related to the creation of models that are going to predict the numerical response variable 'fund_return_10years'.
```

# Null Model
```{r}
null_model_1 <- mean(train$fund_return_10years, na.rm=TRUE)
results_null_1 <- data.frame(pred = null_model_1, original = test$fund_return_10years)
results_null_1 <- results_null_1 %>% 
  mutate(resid = round(pred - original, 2))
cat("Sum of residuals squared of the null model: ", sum(results_null_1$resid^2))
cat("\nMean of the null model: ", null_model_1)
```

# Linear Regression Models - Training
# Linear Model 1 - Training
```{r}
lm1_train <- lm(fund_return_10years ~ return_rating + price_earnings_ratio + price_book_ratio + asset_cash + asset_stocks + asset_bonds + sector_basic_materials + sector_utilities + sector_communication_services + sector_technology + fund_return_ytd + category_return_1year, data = train)

msummary(lm1_train)
```
# Linear Model 2 - Training
```{r}
lm2_train <- lm(fund_return_10years ~ return_rating + price_book_ratio + asset_cash + asset_stocks + sector_basic_materials + sector_utilities + sector_technology + category_return_1year, data = train)

msummary(lm2_train)
```

# Predictive Analysis for Linear Model 1
```{r}
pred_lm1 <- predict(lm1_train, newdata=test)
results_lm1 <- data.frame(pred = pred_lm1, original = test$fund_return_10years)
results_lm1 <- results_lm1 %>% 
  mutate(resid = round(pred - original, 2))
cat("Sum of residuals squared for Linear Model 1:", sum(results_lm1$resid^2))
```

# Predictive Analysis for Linear Model 2
```{r}
pred_lm2 <- predict(lm2_train, newdata=test)
results_lm2 <- data.frame(pred = pred_lm2, original = test$fund_return_10years)
results_lm2 <- results_lm2 %>% 
  mutate(resid = round(pred - original, 2))
cat("Sum of residuals squared for Linear Model 2:", sum(results_lm2$resid^2))
```

# Regression Tree Models - Training
# Regression Tree Model 1 - Training
```{r}
reg_form_1 <- as.formula("fund_return_10years ~ return_rating + price_book_ratio + asset_cash + asset_stocks + sector_basic_materials + sector_utilities + sector_technology + category_return_1year")
regtree_1 <- rpart(reg_form_1, data=train)
regtree_1
```

# Regression Tree Model 1 - Predictive Analysis
```{r}
pred_tree_1 = predict(regtree_1, type="vector")
rsq_1 <- function (x, y) cor(x, y) ^ 2
cat("R-Squared for Regression Tree Model 1:", rsq_1(train$fund_return_10years, pred_tree_1))

results_tree_1 <- data.frame(pred = pred_tree_1, original = train$fund_return_10years)
results_tree_1 <- results_tree_1 %>% 
  mutate(resid = round(results_tree_1$pred - results_tree_1$original, 2))
cat("\nSum of residuals squared for Regression Tree Model 1:", sum(results_tree_1$resid^2))
```

# Regression Tree Model 2 - Training
```{r}
reg_form_2 <- as.formula("fund_return_10years ~ return_rating + asset_stocks + sector_basic_materials + sector_technology + category_return_1year")
regtree_2 <- rpart(reg_form_2, data=train)
regtree_2
```

# Regression Tree Model 2 - Predictive Analysis
```{r}
pred_tree_2 = predict(regtree_2, type="vector")
rsq_2 <- function (x, y) cor(x, y) ^ 2
cat("R-Squared for Regression Tree Model 2: ", rsq_2(train$fund_return_10years, pred_tree_2))

results_tree_2 <- data.frame(pred = pred_tree_2, original = train$fund_return_10years)
results_tree_2 <- results_tree_2 %>% 
  mutate(resid = round(results_tree_2$pred - results_tree_2$original, 2))
cat("\nSum of residuals squared for Regression Tree Model 2:", sum(results_tree_2$resid^2))
```

# Random Forest Models - Training
# Random Forest Model 1 - Training
```{r}
forest_form_1 <- as.formula("fund_return_10years ~ return_rating + price_book_ratio + asset_cash + asset_stocks + sector_basic_materials + sector_utilities + sector_technology + category_return_1year")
forest_1 <- randomForest(forest_form_1, data=train, ntree=201, mtry=3)
forest_1
```

# Random Forest Model 1 - Predictive Analysis
```{r}
pred_forest_1 <- predict(forest_1, newdata = test) 
mse_forest_1 <- sum((pred_forest_1 - test$fund_return_10years)^2/nrow(test))
cat("MSE: ", mse_forest_1)
```

# Random Forest Model 2 - Training
```{r}
forest_form_2 <- as.formula("fund_return_10years ~ return_rating + price_book_ratio + asset_cash + asset_stocks + sector_basic_materials + sector_utilities + sector_technology + category_return_1year")
forest_2 <- randomForest(forest_form_1, data=train, ntree=201, mtry=5)
forest_2
```

# Random Forest Model 2 - Predictive Analysis
```{r}
pred_forest_2 <- predict(forest_2, newdata = test) 
mse_forest_2 <- sum((pred_forest_2 - test$fund_return_10years)^2/nrow(test))
cat("MSE: ", mse_forest_2)
```

# Random Forest Model 3 - Training
```{r}
forest_form_3 <- as.formula("fund_return_10years ~ return_rating + price_book_ratio + asset_cash + asset_stocks + sector_basic_materials + sector_utilities + sector_technology + category_return_1year")
forest_3 <- randomForest(forest_form_3, data=train, ntree=501, mtry=3)
forest_3
```

# Random Forest Model 3 - Predictive Analysis
```{r}
pred_forest_3 <- predict(forest_3, newdata = test) 
mse_forest_3 <- sum((pred_forest_3 - test$fund_return_10years)^2/nrow(test))
cat("MSE: ", mse_forest_3)
```

# Random Forest Model 4 - Training
```{r}
forest_form_4 <- as.formula("fund_return_10years ~ price_book_ratio + asset_cash + asset_stocks + sector_technology + category_return_1year")
forest_4 <- randomForest(forest_form_4, data=train, ntree=201, mtry=5)
forest_4
```


```{r}
pred_forest_4 <- predict(forest_4, newdata = test) 
mse_forest_4 <- sum((pred_forest_4 - test$fund_return_10years)^2/nrow(test))
cat("MSE: ", mse_forest_4)
```

```{r}
#rating, return_rating, investment_type_encoded, size_type_encoded, price_earnings_ratio, price_book_ratio, price_sales_ratio, asset_cash, asset_stocks, asset_bonds, sector_basic_materials, sector_basic_utilities, sector_communication_services, sector_technology, fund_return_ytd, fund_return_1year, category_return_1year, fund_return_3years, category_return_3years, fund_return_5years, category_return_5years, fund_beta_3years, category_beta_3years, fund_beta_5years, category_beta_5years, fund_mean_annual_return_3years, fund_mean_annual_return_3years, net_asset_value
```


# Neural Network Models - Training
# Neural Network Model 1 - Training
```{r}
neural_form_1 <- as.formula("fund_return_10years ~ return_rating + price_book_ratio + asset_cash + asset_stocks + sector_technology")

nn_1 <- neuralnet(neural_form_1, data = train, hidden=c(3,2), linear.output = FALSE)
plot(nn_1)
```

```{r}
t <- test %>% 
  dplyr::select(fund_return_10years, return_rating ,price_book_ratio , asset_cash, asset_stocks,sector_technology, net_asset_value,category_return_3years, fund_return_5years, category_return_5years, investment_type_encoded, size_type_encoded)
```

# Neural Network Model 1 - Predictive Analysis
```{r}
pred_nn_1<- compute(nn_1, t)
pred_nn_result_1 <- pred_nn_1$net.result*(max(funds_scaled$fund_return_10years) - min(funds_scaled$fund_return_10years))+min(funds_scaled$fund_return_10years)

MSE.nn <- sum((t - pred_nn_result_1)^2)/nrow(t)
cat("MSE for NN:", MSE.nn, "\n")
plot(t$fund_return_10years, pred_nn_result_1, col='blue', pch=16, ylab = "predicted rating NN", xlab = "real rating")
```

```{r}
#Test the resulting output
results1 <- data.frame(actual1 = t$fund_return_10years, prediction1 = pred_nn_1$net.result)
roundedresults1 <-sapply(results1,round,digits=0)
roundedresultsdf1 =data.frame(roundedresults1)
table_nn1 <- table(roundedresultsdf1$actual1,roundedresultsdf1$prediction1)
cat("\nAccuracy:", sum(diag(table_nn1))/nrow(t))
```


# Neural Network Model 2 - Training
```{r}
neural_form_2 <- as.formula("fund_return_10years ~ return_rating + price_book_ratio + asset_cash + asset_stocks + sector_technology")
nn_2 <- neuralnet(neural_form_2, data = train, hidden=c(5,3,2),linear.output = FALSE)
plot(nn_2)
```

# Neural Network Model 2 - Predictive Analysis
```{r}
pr_nn_2 <- compute(nn_2, t)
nn_results_2 <- pr_nn_2$net.result*(max(funds_scaled$fund_return_10years) - min(funds_scaled$fund_return_10years))+min(funds_scaled$fund_return_10years)
MSE.nn2 <- sum((t - nn_results_2)^2)/nrow(t)
cat("MSE for NN:", MSE.nn2)
plot(t$fund_return_10years, nn_results_2, col='blue', pch=16, ylab = "predicted rating NN", xlab = "real rating")
```

```{r}
#Test the resulting output
results2 <- data.frame(actual2 = t$fund_return_10years, prediction2 = pr_nn_2$net.result)
roundedresults2 <-sapply(results2,round,digits=0)
roundedresultsdf2 = data.frame(roundedresults2)

table_nn2 <- table(roundedresultsdf2$actual2,roundedresultsdf2$prediction2)
cat("\nAccuracy:", sum(diag(table_nn2))/nrow(t))
```

# Categorical response variable -> mutate -> return_over_inflation
```{r}
Funds2 <- new_funds %>% 
  dplyr::select(rating, return_rating, risk_rating, investment_type_encoded, size_type_encoded, price_earnings_ratio, price_book_ratio, price_sales_ratio, asset_cash, asset_stocks, asset_bonds, asset_others, asset_preferred, asset_convertable, sector_basic_materials, sector_consumer_cyclical, sector_financial_services, sector_real_estate, sector_consumer_defensive, sector_healthcare, sector_utilities, sector_communication_services, sector_energy, sector_industrials, sector_technology, fund_return_ytd, fund_return_1year, category_return_1year, fund_return_3years, category_return_3years, fund_return_5years, category_return_5years, fund_return_10years, category_return_10years, fund_beta_3years, category_beta_3years, fund_beta_5years, category_beta_5years, fund_beta_10years, category_beta_10years, fund_mean_annual_return_3years, fund_mean_annual_return_3years, fund_mean_annual_return_10years, net_asset_value)
```

# Categorical response variable training/testing data sets
```{r}
Funds2 <- na.omit(Funds2)
Funds2 <- Funds2 %>% 
  mutate(Funds2, return_over_inflation=ifelse(fund_return_10years >= 12.37, TRUE, FALSE))
Funds2
```

# Performs standardization
```{r}
# Remove rows that contain missing data.
Funds2 <- Funds2[complete.cases(Funds2),]
Funds2

# Convert columns to numeric
cols <- colnames(Funds2)
Funds2[cols] <- sapply(Funds2[cols],as.numeric)
sapply(Funds2, class)

# Max and mins for standardization
maxs <- apply(Funds2, 2, max) 
mins <- apply(Funds2, 2, min)

# Perform scaling
funds_scaled_2 <- as.data.frame(scale(Funds2, center = mins, scale = maxs - mins))
head(funds_scaled_2, 10) # Print scaled data set
```

```{r}
set.seed(200)
n <- nrow(funds_scaled_2)
test_idx <- sample.int(n, size=round(0.2 * n))
train2 <- funds_scaled_2[-test_idx,]
test2 <- funds_scaled_2[test_idx,]
```

# Null Model
```{r}
null_model_2 <- prop.table(table(Funds2$return_over_inflation))
null_model_2
```
# Multiple Regression Models
# Multiple Logistic Regression Model 1 - Training
```{r}
lg1_train <- glm(return_over_inflation ~ net_asset_value + return_rating + asset_cash + asset_stocks + sector_technology + size_type_encoded, data = train2, family = binomial)
summary(lg1_train)
```

# Multiple Logistic Regression Model 1 - Predictive Analysis
```{r}
# Let's take a look at how the model performs with predictions.
logistic_1_probs = predict(lg1_train, type="response")
favstats(logistic_1_probs)

# Build predictions based on probabilities.
logistic_1_pred = rep(FALSE, nrow(train2))
logistic_1_pred[logistic_1_probs > 0.5] = TRUE

# Misclassification table and accuracy.
table1 <- table(logistic_1_pred, train2$return_over_inflation)
table1
cat("\nAccuracy: ", sum(diag(table1))/sum(table1))
```

# Multiple Logistic Regression Model 2 - Training
```{r}
lg2_train <- glm(return_over_inflation ~ net_asset_value + return_rating + asset_stocks + sector_technology + size_type_encoded, data = train2, family = binomial)
summary(lg2_train)
```

# Multiple Logistic Regression Model 2 - Predictive Analysis
```{r}
# Let's take a look at how the model performs with predictions.
logistic_2_probs = predict(lg2_train, type="response")
favstats(logistic_2_probs)

# Build predictions based on probabilities.
logistic_2_pred = rep(FALSE, nrow(train2))
logistic_2_pred[logistic_2_probs > 0.5] = TRUE

# Misclassification table and accuracy.
table2 <- table(logistic_2_pred, train2$return_over_inflation)
table2
cat("\nAccuracy: ", sum(diag(table2))/sum(table2))
```

# Classification Tree Models
# Classification Tree Model 1 - Training
```{r}
class_tree_form <- as.formula("return_over_inflation ~ net_asset_value + return_rating + asset_cash + asset_stocks + sector_technology + size_type_encoded")
class_tree_1 <- rpart(class_tree_form, data=train2)
class_tree_1
```

# Classification Tree Model 1 - Predictive Analysis
```{r}
test21 <- test2 %>%
   dplyr::mutate(x_return = ifelse(predict(class_tree_1, newdata = test2) >= 0.5, 1, 0))
confusion <- tally(x_return ~ return_over_inflation, data=test21, format="count")
cat("Accuracy:", sum(diag(confusion))/nrow(test21))
```

# Classification Tree Model 2 - Training
```{r}
class_tree_form_2 <- as.formula("return_over_inflation ~ net_asset_value + return_rating + asset_cash + sector_technology + asset_stocks")
class_tree_2 <- rpart(class_tree_form_2, data=train2)
class_tree_2
```

# Classification Tree Model 2 - Predictive Analysis
```{r}
test22 <- test2 %>%
   dplyr::mutate(x_return = ifelse(predict(class_tree_2, newdata = test2) >= 0.5, 1, 0))
confusion2 <- tally(x_return ~ return_over_inflation, data=test22, format="count")
cat("Accuracy:", sum(diag(confusion2))/nrow(test22))
```

# Random Forest Models
# Random Forest Model 1 - Training
```{r}
train2$return_over_inflation <- as.character(train2$return_over_inflation)
train2$return_over_inflation <- as.factor(train2$return_over_inflation)

f <- as.formula("return_over_inflation ~ net_asset_value + return_rating + asset_cash + asset_stocks + sector_technology + size_type_encoded")
mod_forest <- randomForest(f, data=train2, ntree=200, mtry=3)
mod_forest
```

# Random Forest Model 1 - Predictive Analysis
```{r}
sum(diag(mod_forest$confusion))/nrow(train2)
```

# Random Forest Model 2 - Training
```{r}
train2$return_over_inflation <- as.character(train2$return_over_inflation)
train2$return_over_inflation <- as.factor(train2$return_over_inflation)
f2 <- as.formula("return_over_inflation ~ net_asset_value + return_rating + asset_cash + asset_stocks + sector_technology")
mod_forest2 <- randomForest(f2, data=train2, ntree=200, mtry=5)
mod_forest2
```

# Random Forest Model 2 - Predictive Analysis
```{r}
sum(diag(mod_forest2$confusion))/nrow(train2)
```

# Random Forest Model 3 - Training
```{r}
train2$return_over_inflation <- as.character(train2$return_over_inflation)
train2$return_over_inflation <- as.factor(train2$return_over_inflation)
f3 <- as.formula("return_over_inflation ~ net_asset_value + return_rating + asset_cash + asset_stocks + sector_technology")
mod_forest3 <- randomForest(f3, data=train2, ntree=300, mtry=3)
mod_forest3
```

# Random Forest Model 3 - Predictive Analysis
```{r}
sum(diag(mod_forest3$confusion))/nrow(train2)
```

# KNN Models
# KNN Model 1 - Training and Predictive Analysis
```{r}
train3 <- train2 %>% 
  dplyr::select(return_rating , sector_technology , net_asset_value, asset_cash, asset_stocks)

test3 <- test2 %>% 
  dplyr::select(return_rating , sector_technology , net_asset_value, asset_cash, asset_stocks)

knn_1 <- knn(train=train3, test=test3, cl=train2$return_over_inflation, k=2)
confusion_knn <- table(knn_1, test2$return_over_inflation)
confusionMatrix(confusion_knn)
```


# KNN Model 2 - Training and Predictive Analysis
```{r}
knn_2 <- knn(train=train3, test = test3, cl=train2$return_over_inflation, k=4)
confusion_knn2 <- table(knn_2, test2$return_over_inflation)
confusionMatrix(confusion_knn2)
```

# Naive Bayes Models
# Naive Bayes Model 1 - Training and Predictive Analysis
```{r}
f <- as.formula("return_over_inflation ~ net_asset_value + return_rating + asset_cash + asset_stocks + sector_technology + size_type_encoded")

nb_1 <- naiveBayes(f, data=train2)
nb_pred_1 <- predict(nb_1, newdata=train2)
confusiontable<- table(nb_pred_1, train2$return_over_inflation)
confusionMatrix(confusiontable)
```

# Naive Bayes Model 2 - Training and Predictive Analysis
```{r}
f2 <- as.formula("return_over_inflation ~ net_asset_value + return_rating + sector_technology + size_type_encoded")

nb_2 <- naiveBayes(f2, data=train2)
nb_pred_2 <- predict(nb_2, newdata=train2)
confusiontable2 <- table(nb_pred_2, train2$return_over_inflation)
confusionMatrix(confusiontable2)
```